Automatic Singing Transcription: Phonemes and Alignment

### 2023-12-05 report/paper.tex
add future works

\subsubsection{Transforming Utterance level datasets into Phoneme level datasets} \label{sec:utteranceToPhoneme}
A technique from ASR that can also be applied to AST is transforming utterance/sentence
level datasets into phoneme level datasets. There are many ASR datasets that contain single utterances
such as LibriSpeech and JamendoLyrics. These datasets can be effectively
transformed into phoneme level datasets by using a phoneme dictionary such as the CMU Pronouncing
Dictionary \citep{CMUDict}. This dictionary contains a mapping from words to phonemes. Using this
dictionary, the utterances can be transformed into phoneme sequences. These sequences can then be
used with a Connectionist Temporal Classification (CTC) loss function to train AST models \citep{CTC}.
\begin{figure*}
    \centering
    \includegraphics[width=0.5\textwidth]{assets/CTC.png}
    \caption{Example of CTC Function collapsing repetitions. Image from \citep{CTC}}
    \label{fig:CTC}
\end{figure*}
CTC allows for the model to output a sequence of phonemes per time step and the
repetitions are collapsed into a single phoneme as shown in Figure \ref{fig:CTC}.
This technique was used to allow a model to incorperate the utterance level LibriSpeech
dataset into the training for a phoneme ASR model \citep{wav2vec}. Timestamps can be
retrieved from the pre-CTC output that had time-aligned phoneme classifications.
The same process can be applied to AST models to allow them to use the utterance level song datasets.

@@

\subsection{Existing Datasets}

@@

@@

@@

% \subsection{VAD Performance}

@@

\section{Future Work}
\subsection{Train on More Datasets}
\subsection{Finetuning Wav2Vec 2.0 Feature Extractor}
\subsection{Adversarial Training}
\subsection{Transform Singing to Speech Data}
remove pitch from singing
\subsection{Phoneme to Grapheme}
\citep{G2P} or \citep{P2G}

>>>>

\subsubsection{Transforming Utterance level datasets into Phoneme level datasets} \label{sec:utteranceToPhoneme}
A technique from ASR that can also be applied to AST is transforming utterance/sentence
level datasets into phoneme level datasets. There are many ASR datasets that contain single utterances
such as LibriSpeech and JamendoLyrics. These datasets can be effectively
transformed into phoneme level datasets by using a phoneme dictionary such as the CMU Pronouncing
Dictionary \citep{CMUDict}. This dictionary contains a mapping from words to
ARPAbet style phonemes \citep{CMUDict}. Using this
dictionary, the utterances can be transformed into phoneme sequences. These sequences can then be
used with a Connectionist Temporal Classification (CTC) loss function to train AST models \citep{CTC}.
\begin{figure*}
    \centering
    \includegraphics[width=0.5\textwidth]{assets/CTC.png}
    \caption{Example of CTC Function collapsing repetitions. Image from \citep{CTC}}
    \label{fig:CTC}
\end{figure*}
CTC allows for the model to output a sequence of phonemes per time step and the
repetitions are collapsed into a single phoneme as shown in Figure \ref{fig:CTC}.
This technique was used to allow a model to incorperate the utterance level LibriSpeech
dataset into the training for a phoneme ASR model \citep{wav2vec}. Timestamps can be
retrieved from the pre-CTC output that had time-aligned phoneme classifications.
The same process can be applied to AST models to allow them to use the utterance level song datasets.

@@

\subsection{Existing Datasets} \label{sec:datasets}

@@

In order to transform these datsets into a valid input for the model, datasets that only contain
utterances will be transformed into phoneme level datasets using the CMU Pronouncing Dictionary
(CMUDict) and then tokenized. Luckily, the TIMIT dataset already uses an extended version of the
ARPAbet phonemes so they can be remapped into a smaller subset of phonemes. The CMUDict also
includes stresses in the form of numbers at the end of some phonemes that were removed. Finally,
the unknown and padding tokens were added to the phoneme vocabulary. Any words that appeared in
the datasets but were not in CMUDict were manually added to the dictionary. HuggingFace and PyTorch
constructs to help with this process and allow for batching.

@@

\subsection{Base Model}
In order to get a baseline for our improvements, we train a base model on the TIMIT dataset. This
This dataset is a one of the most popular and defacto standard for ASR models that do phoneme
recognition. The starting model that will be finetuned on TIMIT will be the `facebook/wav2vec2-base-960h'
model from the HuggingFace \citep{huggingFaceWav2vec2}.

\subsection{Singer Model}

@@

% \subsection{VAD Performance}
\subsection{Phoneme Recognition Performance}

@@

\section{Future Work}
\subsection{Train on More Datasets}
Dispite the lack of datasets, there are still many datasets listed in section \ref{sec:datasets}
that can be used to train AST models. Deep learning networks are known to perform better with more
training data and the greater variety of data will allow the model to generalize better.
datasets used in the training of our models were TIMIT and CSD. However, there are many genres of
music that are not represented in children's songs such as rap.

\subsection{Finetuning Wav2Vec 2.0 Feature Extractor}
The FMA dataset described in section \ref{sec:datasets} can be used to finetune the wav2vec 2.0
feature extractor. This will allow the model to learn features that are more specific to singing
and music. This dataset will need some extra preprocessing to detect singing because the songs
include music with no vocals. This can be done by using a VAD model to detect the singing and
removing the parts of the audio that do not contain singing. The wav2vec 2.0 paper mentions how
the model goes through self-supervised learning to learn discrete speech units \citep{wav2vec} and
a large dataset like FMA will allow the model to learn discrete singing units.

\subsection{Generative Adversarial Networks}
Generative Adversarial Networks (GANs) are a type of neural network that can be used to generate
new data and can have recently become popular with the success of Midjourney and
Stable Diffusion. Voice synthesis and conversion have also improved with services that
clone voices or text-to-speech services. The intersection of these technologies can
be used to generate new songs. The songs that are generated do not need to be top highlights
but plausible sounding songs that can be used to train AST models. Using a GAN can generate
songs with more variety than the pitch shifted songs mentioned in \ref{sec:pseudoSinging} and
can lead to better generalization.

\subsection{Phoneme to Grapheme}
One of the main applications of AST is to generate and sync lyrics to songs for karaoke
applications. However, most people cannot read ARPAbet phonemes or IPA phonemes. Instead,
they can read words and graphemes that form words. Phonemes and graphemes represent the same
unit of sound but the graphemes are more readable because they are just chunks of the original word.
This means that as the words are sung, each grapheme of the word can be highlighted individually.
To get a grapheme output, the phoneme output can be converted to a grapheme output using a
separate phoneme to grapheme model or generate character level output and align the characters and
phonemes using the timestamps.
